Towards Semantic Patches

### 2010-06-18 papers/MIPS2010/abstract.tex
Peaufinage de la description du kernel
git-svn-id: http://gasp.dev.regis-gianas.com/svn/trunk@109 403e655e-980a-47bd-8c60-1ee5348bb6b7

\remplan{On rentre dans les détails techniques. Je crois qu'il faut
  prendre un peu de temps pour définir précisément les termes utilisés
  (repository, patch, etc) et expliquer/paraphraser les règles.}

@@

% Y: Remarque en passant: pour maintenir des notations cohérentes, il
% vaut mieux se définir des macros pour absolument tout
% (metavariables, jugements, ...). Ca passe mieux à l'échelle.  On
% verra ca pour le prochain papier :-).

@@

Following the standard syntax, we denote by $(x:t)\cdot u$ the
dependent product (elsewhere written with a $\Pi$). The new construct
$(x=a:t)\cdot u$ is a definition, whose value is $a$. We write as
usual $(x_1:t_1)(x_2:t_2)\ldots(x_n:t_n)\cdot t$ for
$(x_1:t_1)\cdot(x_2:t_2)\cdot\ldots\cdot(x_n:t_n)\cdot t$, $t \to u$
for $(x:t)\cdot u$ where $x\notin u$, and $(x\ y : t)\cdot u$ for
$(x:t)(y:t)\cdot u$. The lookup in the environment is written either
$\Gamma(x):t$, denoting ``$\Gamma$ contains a declaration $[x:t]$
\emph{or} a definition $[x=a:t]$'', or $\Gamma(x)=a:t$ denoting
``$\Gamma$ contains a definition $[x=a:t]$''. Capture-avoiding
renaming of variables in a term $t$ is written $t\subst{x}{y}$ for
``replace all occurences of $x$ by $y$'' and is defined the usual way.

\paragraph{Typing} The typing of our system is a restriction of the
usual typing rules for dependently-typed systems with the additional
two rules for typing a singleton product and the application of a
singleton argument (featuring the syntactic verification in the
environment). It depends on: 
\begin{itemize}
  \item$\mathcal S$ a set of sorts \sort,
  \item$\mathcal A \in \mathcal S^2$ the axioms of the system,
  \item$\mathcal R \in \mathcal S^3$ the allowed products.
\end{itemize}
It relies on two distinct judgments:
\begin{itemize}
\item $\Gamma\vdash a : t$ ($a$ has type $t$ in $\Gamma$),
\item $\Gamma\vdash t : \sort$ ($t$ has sort $s$ in $\Gamma$).
\end{itemize}
The rules are:
\begin{mathpar}
  \infer[Axiom]{
  }{\Gamma\vdash\sort_1:\sort_2}
  \quad(\sort_1,\sort_2)\in\mathcal A
  \and
  \infer[Init]{
    \Gamma(x):t
  }{\Gamma\vdash x:t}
  \and
  \infer[Prod]{\Gamma\vdash t:\sort_1 \and 
    \Gamma[x:t]\vdash u:\sort_2
  }{\Gamma\vdash(x:t)\cdot u : \sort_3}
  \quad(\sort_1,\sort_2,\sort_3)\in\mathcal R
  \and
  \infer[Sing]{\Gamma\vdash a:t \and
    \Gamma\vdash t:\sort_1 \and
    \Gamma[x=a:t]\vdash u:\sort_2
  }{\Gamma\vdash(x=a:t)\cdot u : \sort_3} 
  \quad(\sort_1,\sort_2,\sort_3)\in\mathcal R
  \and
  \infer[App]{
    \Gamma\vdash a:(y:t)\cdot u \and \Gamma(x):t
    }{\Gamma\vdash a\ x : u\subst{x}{y}}
  \and
  \infer[Skip]{
    \Gamma\vdash a:(y=b:t)\cdot u \and \Gamma(x)=b:t
    }{\Gamma\vdash a : u\subst{x}{y}}
\end{mathpar}

A term $t$ represents a state of the repository at a given moment. It
contains the declaration of the object language's syntax, its typing
rules, the syntactic objects and derivations recorded in it along with
the patches that let the repository evolve. If $\cdot\vdash t:\sort$,
then the repository is well-typed.

>>>>

We describe the state of a repository at a given moment by a type in
the following syntax. Well-typed types in the repository meta-language
guarantee that it contains only well-typed proofs in the object
language.

@@

@@

Following the standard syntax, we denote by $(x:t)\cdot u$ the
dependent product (elsewhere written with a $\Pi$). The new construct
$(x=a:t)\cdot u$ is a definition binder, whose value is $a$. We write
as usual $(x_1:t_1)(x_2:t_2)\ldots(x_n:t_n)\cdot t$ for
$(x_1:t_1)\cdot(x_2:t_2)\cdot\ldots\cdot(x_n:t_n)\cdot t$, $t \to u$
for $(x:t)\cdot u$ where $x\notin u$, and $(x\ y : t)\cdot u$ for
$(x:t)(y:t)\cdot u$. The lookup in the environment is written either
$\Gamma(x):t$, denoting ``$\Gamma$ contains a declaration $[x:t]$
\emph{or} a definition $[x=a:t]$'', or $\Gamma(x)=a:t$ denoting
``$\Gamma$ contains a definition $[x=a:t]$''. Capture-avoiding
renaming of variables in a term $t$ is written $t\subst{x}{y}$ for
``replace all occurences of $x$ by $y$'' and is defined the usual way.

\paragraph{Typing} Typing is parameterized by:
\begin{itemize}
  \item$\mathcal S$ a set of sorts \sort,
  \item$\mathcal A \in \mathcal S^2$ the axioms of the system,
  \item$\mathcal R \in \mathcal S^3$ the allowed products.
\end{itemize}
It relies on two distinct judgments:
\begin{itemize}
\item $\Gamma\vdash a : t$ ($a$ has type $t$ in $\Gamma$),
\item $\Gamma\vdash t : \sort$ (the repository $t$ is well-typed of
  sort $s$ in $\Gamma$).
\end{itemize}
The rules are:
\begin{mathpar}
  \infer[Axiom]{
  }{\Gamma\vdash\sort_1:\sort_2}
  \quad(\sort_1,\sort_2)\in\mathcal A
  \and
  \infer[Init]{
    \Gamma(x):t
  }{\Gamma\vdash x:t}
  \and
  \infer[Prod]{\Gamma\vdash t:\sort_1 \and 
    \Gamma[x:t]\vdash u:\sort_2
  }{\Gamma\vdash(x:t)\cdot u : \sort_3}
  \quad(\sort_1,\sort_2,\sort_3)\in\mathcal R
  \and
  \infer[Let1]{\Gamma\vdash a:t \and
    \Gamma\vdash t:\sort_1 \and
    \Gamma[x=a:t]\vdash u:\sort_2
  }{\Gamma\vdash(x=a:t)\cdot u : \sort_3} 
  \quad (x=a:t)\notin\Gamma
  \and
  \infer[Let2]{\Gamma\vdash t:\sort_1 \and
    \Gamma\vdash u:\sort_2
  }{\Gamma\vdash(x=a:t)\cdot u : \sort_3} 
  \quad(x=a:t)\in\Gamma
  \and
  \infer[App]{
    \Gamma\vdash a:(y:t)\cdot u \and \Gamma(x):t
    }{\Gamma\vdash a\ x : u\subst{x}{y}}
  \and
  \infer[Skip1]{
    \Gamma\vdash a:(y=b:t)\cdot u \and \Gamma(x)=b:t
    }{\Gamma\vdash a : u\subst{x}{y}} \and
  \infer[Skip2]{
    \Gamma[x=b:t]\vdash a:(y=b:t)\cdot u
    }{\Gamma\vdash a : u\subst{x}{y}}
\end{mathpar}

The first four rules are standard. As application is only performed on
variables, the \textsc{App} rule looks up directly in the
environment. Rules \textsc{Let1} and \textsc{Skip1} show the
verification and the use of the equality binder in case the definition
is not already known. Rule \textsc{Let2} and \textsc{Skip2} allow
incremental typing in our system: if the definition was already known
in the environment, we do not need to recheck it. If $\cdot\vdash
t:\sort$, then the repository $t$ is well-typed.
